### 2021-07-05 paper/snippets.tex
first edit

\vskip 6pt
In practice, that means that you, the author of the unit test, has to choose a few possible values for each function argument. These values will be samples from equivalence classes for that argument and may include corner cases. Then the \utd library generates a set of test cases. In the unit testing framework, the code loops over the test cases, checking whether the results are acceptable.